Geometrization of characters of tori over non-Archimedean local fields

### 2013-08-08 shanghai.tex
Going home.

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals formed by applying the functor $\sheafHom(\ - \ ,(\EE)_{\bG})$ to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from second condition appearing in the definition of character
sheaves). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=30, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{t_{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about pullbacks of sheaves and maps of sheaves.
  \David{Is this explanation of $(f\circ g)^* = g^* \circ f^*$ sufficient?}
\end{proof}

@@

Let $\CS_0(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CS_f(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$.

@@

@@

  It is also a matter of routine to verify that every morphism of group schemes $G'\to G$ 
  induces a monoidal functor $C(G) \to C(G')$ by base-change along $G'\to G$ ({\it cf.}, Proposition~\ref{prop:pullback}); 
  indeed, the pull-back of $(f,\psi)$ along any morphism $m : G'\to G$ of group schemes 
  can be read off the short exact sequence obtained by pulling back along $m$. 
  Taking the case when $m$ is the multiplication morphism $m : G\times G\to G$, 
  and using the tensor category structure above, 
  we see that the pull-back of $(f,\psi)$ along $m$ is canonically isomorphic, in $C_1(G\times G)$, to $(f,\psi)\boxtimes(f,\psi)$.

 \Clifton{Insert diagram here.}

@@

 Since $L : C(G) \to \Loc(G)$ takes weak isomorphisms to isomorphisms,
 it factors (uniquely!) through the localisation functor $C(G) \to C(G)[W^{-1}]$ 
 to define a faithful functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$. 
 Since $W$ is saturated, the functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$ is full. 
 %
 Finally, we are ready to exhibit the full functor $L_G: \CS_f(G) \to \CS_0(G)$ promised above. 
 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $C_1(G\times G)$, canonically;
 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that 
  \[
  L_G  \ceq  L_{W}\vert_{\CS_f(G)} : \CS_b(G) \to \CS_0(G) 
  \qquad\text{defined by}\qquad 
  (f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))
  \]
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

\Clifton{I'll address this point in the next section: "Because $\Gal(\bFq/\Fq)$ acts trivially on $\Aut({\bar f})$, the local system
 ${\bar L}_G(f,\psi)\ceq b_G^* L(f,\psi)$ comes equipped with a canonical isomorphism 
 $F : \Frob_G^* {\bar L}_G(f,\psi) \to {\bar L}_G(f,\psi)$ of constructible sheaves on ${\bar G}$."}

@@

\begin{proposition}\label{prop:connected-1}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CS_f(G) \hookrightarrow \CS_0(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  The local system $L(f,\psi)$ (defined in Section~\ref{sec:finite}) naturally

  To see that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$ we simply apply the functions--sheaves dictionary

  We will see that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$, 
  where $L_G$ is the functor constructed in Section~\ref{sec:finite}.
  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it will then follow that $L_G$ is an equivalence.

>>>>

%\begin{proposition}\label{tensor}
The category $\CS(G)$ of character sheaves on $G$ is a rigid monoidal
category [insert reference] under
$\cs{L} \otimes \cs{L'} \ceq (\gcs{L}\otimes\gcs{L'}, \mu\otimes\mu', F\otimes F')$
with duals given by applying the functor $\sheafHom(\ - \ ,(\EE)_{\bG})$ (sheaf hom) to
conditions \ref{CS.0}, \ref{CS.1} and \ref{CS.2}.
%\end{proposition}
(The category of characters sheaves on $G$ is not abelian, so $\CS(G)$
is not a tensor category in the sense of \cite{Deligne:tensorielles}.)
%\begin{proof}\end{proof}
%
This rigid monoidal category structure for $\CS(G)$ puts a group
structure on the set $\CSiso{G}$ of isomorphism classes in
$\CS(G)$; we refer to this as the \cdef{Picard group of character
sheaves} on $G$; {\it cf.} {\tt http://ncatlab.org/nlab/show/Picard+group}.
\Clifton{Replace ncatlab with other references.}

@@

Let $\cs{L}$ be a character sheaf on $G$. Every $g\in G(\Fq)$
determines a point $\bg$ on $\bG$ fixed by $\Frob_G$ and
therefore an automorphism $F_{\bg}$ of $\gcs{L}_{\bg}$. Note
that $\Aut_{\EE}(\gcs{L}_{\bg}) \iso \EEx$, canonically (this
follows from the Condition~\ref{CS.1}). The isomorphism $\mu : \bm^* \gcs{L} \to \LxL$
guarantees that the function $t^\Frob_\cs{L} : G(\Fq)\to \EEx$,
defined by $g \mapsto F_{\bg}$, is a group homomorphism;
we refer to $t^\Frob_\cs{L}$ as the \cdef{trace of Frobenius of $\cs{L}$}.
\Clifton{There is no actual trace involved, since $\gcs{L}$ is a rank
  $1$ local system. More generally, trace is needed.}

@@

\begin{proposition}\label{prop:pullback}
  If $f : H\to G$ is a morphism of commutative group schemes locally of finite type over $\Fq$ then
  $(\gcs{L},\mu,F) \mapsto (\bf^*\gcs{L},(\bfxf)^*\mu,\bf^*F)$
  defines a monoidal functor $f^* : \CS(G) \to \CS(H)$ such that
  \[
  \begin{tikzcd}[row sep=20, column sep=30]
   \CSiso{G} \rar{f^*} \arrow[swap]{d}{t_{G}} & \CSiso{H} \dar{t_H} \\
  \Hom_\text{grp}(G(\Fq),\EEx) \rar & \Hom_\text{grp}(H(\Fq),\EEx)
  \end{tikzcd}
  \]
  is a commutative diagram of groups, where the lower homomorphism is
  dual to $f : H(\Fq)\to G(\Fq)$.  Moreover, $(f\circ g)^* = g^* \circ f^*$.
\end{proposition}

@@

  Finally, the fact that $(f\circ g)^* = g^* \circ f^*$ follows from the analogous
  statements about the pullbacks functor on $\ell$-adic constructible sheaves.
 % \David{Is this explanation of $(f\circ g)^* = g^* \circ f^*$ sufficient?}
\end{proof}

@@

Let $\CS_0(G)$ be the category of pairs $(\cs{L}_0,\mu_0)$ 
where $\cs{L}_0$ is a constructible $\ell$-adic sheaf on $G$, 
which is a rank-$1$ local system on each connected component of $G$, 
equipped with an isomorphism $\mu_0 : m^* \cs{L}_0 \to \cs{L}_0 \boxtimes \cs{L}_0$ 
satisfying the analogue of Condition~\ref{CS.1} on $G$; 
morphisms in $\CS_f(G)$ are defined as in (the second part of) Condition~\ref{CS.2}. 
This is the category of \cdef{bounded character sheaves} on $G$. 
This is a rigid monoidal category in the obvious way, 
with duals given by $\cs{L}_0^\vee \ceq \sheafHom(\cs{L}_0,(\EE)_{G})$.

@@

\Clifton{Add the monoidal category structure here.}

@@

@@

 Since $L : C(G) \to \Loc(G)$ takes weak isomorphisms to isomorphisms,
 it factors (uniquely!) through the localisation functor $C(G) \to C(G)[W^{-1}]$ 
 to define a faithful functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$. 
 Since $W$ is saturated, the functor $L_{W} : C(G)[W^{-1}] \to \Loc(G)$ is full.

 One more ingredient is needed before we can exhibit the full functor $L_G: \CS_f(G) \to \CS_0(G)$ promised above: 
 we must put rigid monoidal category structures on $\CS_f(G)$ and $\CS_0(G)$.

  It is also a matter of routine to verify that every morphism of group schemes $G'\to G$ 
  induces a monoidal functor $C(G) \to C(G')$ by base-change along $G'\to G$ ({\it cf.}, Proposition~\ref{prop:pullback}); 
  indeed, the pull-back of $(f,\psi)$ along any morphism $m : G'\to G$ of group schemes 
  can be read off the short exact sequence obtained by pulling back along $m$. 
  Taking the case when $m$ is the multiplication morphism $m : G\times G\to G$, 
  and using the tensor category structure above, 
  we see that the pull-back of $(f,\psi)$ along $m$ is canonically isomorphic, in $C_1(G\times G)$, to $(f,\psi)\boxtimes(f,\psi)$.
  \Clifton{I need to work on this diagram and explain better what's going on here. Under construction.}
  \[
   \begin{tikzcd}
V & \Aut(V) & \arrow[swap]{l}{\psi} \ker f \arrow{r}  & H \arrow{r}{f} & G \\
V & \Aut(V) \arrow[equal]{u} & \arrow[swap]{l}{\psi_{m_H}} \ker (f\times f) \arrow{r} \arrow{u}{m_H} & \arrow{d} H\times H \arrow{r}{f\times f} \arrow{u}{m_H} & G\times G \arrow{u}{m} \\
V\otimes V & \Aut(V\otimes V) & \arrow[swap]{l}{\psi\otimes \psi} \ker f \times \ker f \arrow{r}  & H\times H \arrow{r}{f\times f}  & G\times G \arrow[equal]{u} \\
   \end{tikzcd}
  \]

 If $(f,\psi)$ in an object in $C_1(G)$ then, as we have already seen, 
 $m^*L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $C_1(G\times G)$, canonically;
 let $\mu(f,\psi)$ be that isomorphism of $\ell$-adic sheaves on $G$,
 It follows immediately that the functor $L_G  \ceq  L_{W}\vert_{\CS_f(G)} : \CS_b(G) \to \CS_0(G)$, 
 given on objects by $(f,\psi) \mapsto (L(f,\psi), \mu(f,\psi))$,
 is a full and faithfull monoidal functor, completing the proof.
\end{proof}

%\Clifton{I'll address this point in the next section: "Because $\Gal(\bFq/\Fq)$ acts trivially on $\Aut({\bar f})$, the local system ${\bar L}_G(f,\psi)\ceq b_G^* L(f,\psi)$ comes equipped with a canonical isomorphism $F : \Frob_G^* {\bar L}_G(f,\psi) \to {\bar L}_G(f,\psi)$ of constructible sheaves on ${\bar G}$."}

@@

\begin{proposition}\label{prop:connected-1}
  If $G$ is a connected commutative algebraic group over $\Fq$ then 
  the full subcategories $\CS_f(G) \hookrightarrow \CS_0(G) \hookrightarrow \CS(G)$ 
  appearing in Section~\ref{sec:bounded} are equivalences 
  and (with reference to Section~\ref{sec:Frob})
  \[
  t_{G} : \CSiso{G} \to \Hom_\text{grp}(G(\Fq),\EEx)
  \]
  is an isomorphism of groups.
\end{proposition}

@@

  We now show that $L_G(f,\psi) \iso (\cs{L}_0,\mu_0)$. 
  First, recall that $L_G(f,\psi) = (L(f,\psi), \mu(f,\psi))$. 
  By  \cite{Laumon}*{1.1.3.3}, $t_{L(f,\psi)}^{\Frob^n} = t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q}$, for each positive integer $n$
  On the other hand, $t_{\cs{L}}^{\Frob}\circ N_{\FF_{q^n}/\FF_q} = 
  Using notation from \cite{Laumon}*{1.1},  this means $t_{\bullet L(f,\psi)} = t_{\bullet \cs{L}_0}$.
  It now \cite{Laumon}*{Th\'eor\`eme~1.1.2}, it follows that  $L(f,\psi) \iso \cs{L}_0$ in $\Loc(G)$ (admittedly, this is a big hammer).

   we simply apply the functions--sheaves dictionary
  (see \cite{Laumon}*{1.1.3} or \cite{SGA4.5}*{Sommes trigonométriques}) as follows.

  As we saw in Section~\ref{sec:finite}, the $\ell$-adic local system $L(f,\psi)$ on $G$ comes equipped with an isomorphism
  $m^* L(f,\psi) \iso L(f,\psi) \boxtimes L(f,\psi)$ in $\CS_b(G)$; 
  in the case at hand, this isomorphism is \cite[1.1.3.2]{Laumon}.
  By \cite[1.1.3.2]{Laumon},

  Since we have already seen (in Section~\ref{sec:finite}) that $L_G$ is full and faithful, 
  it will then follow that $L_G$ is an equivalence.
